Parametric controls on vegetation responses to biogeochemical forcing in the CLM5

### 2019-07-03 main_revision.tex

At CAX, NWT and HVF, higher $N_{costs}$ increases the responsiveness of the system to N, and to a lesser extent CO$_{2}$.  The magnitude of the impact on N fertilization appeared to be related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).   Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF.

The fraction of N fixers (\emph{frac\_fixers}) has a particularly dominant role over the model response to step changes in N and to CO$_{2}$. For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP_CN}, \ref{LAI_CN}).   At MET,  the changing fixation capacity had no impact on the response of GPP or LAI to N fertilization.  Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization for this site (which had the smallest PNU in the first instance), suggesting fundamental metabolic limitations to carbon economics and growth (e.g. a short growing season).

The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state. The primary impacts were at, an exception being Niwot Ridge, where low values of \emph{leafcn} had a low response to N fertilization These cases suffered from high water limitation, given their high $N\_{leaf}$ values and high gas exchange fluxes (Figure \ref{btran state}) and thus CO$_{2}$ fertilization alleviates this problem. At CAX, high \emph{leafcn} cases had stronger responses to N fertilization, reflecting the higher PNU in those cases prior to fertilizaion.

At Niwot Ridge, runs with low values of the \emph{medlyn\_slope} parameter (and thus high water use efficiency) were able to respond very strongly to increasing Nitrogen (SI Figure 4), given their lower initial water stress (Figure \ref{btran state}). Surprisingly, however, given its critical role in the responsiveness of stomata to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response, potentially suggesting that the impact on the state condition had already accounted for the impact of this parameter.

In general the impacts of the maintenance respiration intercept \emph{lmr\_intercept} were small, and limited to the impact of large values increasing fertilization responses to CO$_{2}$ and N, potentially on account of the low LAI and thus greater feedback between LAI and GPP, and in line with its limited impact on the system state, \emph{frac\_ectomy\_fungi} had no discernible impact on fertilization responses.

There was limited impact of \emph{cn\_flex\_a} on the fertilization responses, as per it's impact on the state.  At Harvard Forest, low values of  \emph{cn\_flex\_b} increased responses to N deposition and decreased response to CO$_{2}$. $N_{leaf}$ followed suit, increasing substantially in the N deposition experiment. Low values of \emph{cn\_flex\_b} decrease the amount of carbon spent on N uptake when N costs are high, therefore potentially suppressing $N_{leaf}$. Adding N overcomes this suppression, leading to greater fertilization responses.

The impact of \emph{cn\_flex\_c} (the parameter controlling the degree to which plant C:N ratios are constrained to their target) appeared to have its strongest impact on the fertilization responses of vegetation carbon (SI Figure 2)  at Harvard Forest and Niwot Ridge (and Caxiuan\~a, to a smaller extent). These impacts do not mirror those on leaf area index or NPP, and overall the response to this parameter appears complex.  At Metolius and Niwot, \emph{cn\_flex\_c} has a dominant control over the impacts of CO$_{2}$ fertilization on $N_{leaf}$ (SI Figure 1).  The impacts on vegetation carbon are likely due to the larger decreases in tissue N content that are permissable when values of this parameter are high (e.g. SI Figure 3).  Low  $N_{leaf}$ values permit higher leaf area and vegetation carbon, but reduce assimilation rates, as illustrated at Harvard Forest and Niwot Ridge (SI Figure 5 \& 7).

@@

To interrogate this further, we ran an entire separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition. In the default parameterization CO$_{2}$ and N fertilization increased LAI, amplifying these magnitude of ecosystem responses to these perturbations. For example, at Harvard Forest, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.18, 1.17 and 1.22 for GPP, NPP and LAI, respectively, in the high LAI case vs. 1.43, 1.66, 1.47 in the control) and a lower N response as well (1.05, 1.20 and 1.15 for GPP, NPP, and LAI, respectively, in the high LAI case vs. 1.1, 1.26, 1.21 for the control). Similar results were found at all other sites.  Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology, and this may explain the large fertilization responses found here compared to those of \cite{wieder2019}.

\section{Discussion}

The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.

At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (Wieder et al. 2018).  The trade-off between responses to N fertilization and CO$_{2}$ response when the fraction of N fixing plants varies mirrors the change in fertilization response between CLM4, CLM4.5 and CLM5, and the CLM5 response is consistant with the capacity of the model to increase fixation rates in response to higher carbon dioxide levels. The number of degrees of freedom illustrated here precludes, however, proof that this is the only contributing factor. This analysis indicates that global experiments with lower or zero \emph{frac\_fixers} would help to ascertain whether this is indeed the case.

@@

  \begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_CNtime_defpft.png}
     \caption{Fertilization responses of the default model though time since the (step-change) in CO$_{2}$ (top row) and Nitrogen deposition (bottom row) for CAX (blue), NWR (red), HVF (yellow), MET (purple).}
     \label{MET state}
 \end{figure}

>>>>

Higher $N_{costs}$ increase the responsiveness of the system to N at CAX, NWT and HVF.  The magnitude of the impact on N fertilization ism as might be expected, related to the extent of N limitation in the pre-fertilization state (Figure \ref{btran state}).   Very low values of $N_{costs}$ appeared to reduce the responsiveness of the system to CO$_{2}$ fertilization, particularly at CAX and HVF **.

\emph{frac\_fixers} (the maximum fraction of NPP available for N fixation) had a particularly dominant role over the model response to step changes in N and to CO$_{2}$. For GPP and LAI, all sites displayed a 'trade-off', whereby high \emph{frac\_fixers} systems responded strongly to CO$_{2}$ but not N, and low \emph{frac\_fixers} systems responded strongly to N but not CO$_{2}$ (Figure \ref{GPP_CN}, \ref{LAI_CN}).   At MET,  the changing fixation capacity had no impact on the response of GPP or LAI to N fertilization.  Only \emph{slatop} and \emph{stem\_leaf} had large impacts on N fertilization for this site (which had the smallest PNU in the first instance), suggesting fundamental metabolic limitations to carbon economics and growth (e.g. a short growing season).

The impact of \emph{leafcn} on fertilization responses was not as dominant a control over fertilization responses as it was over the system state. Complex responses were illustrated, reflecting the differential costs and benefits of Ni on the system.  For N fertilization, at CAX and NWR, lower values of \emph{leafcn} had smaller responses to N fertilization, reflecting the higher initial LAI of these configurations and the non-linear GPP feedback discusses earlier.   At CAX, the highest \emph{leafcn} had smaller CO$_{2}$ responses. (Figure \ref{btran state} illustrates that this configuration had less water stress, given that both the lower LAI and lower photosynthetic capacity should lead to smaller stomatal conductances and evapotranspiration. and thus had less water stress to be alleviated by the high CO$_{2}$.

Surprisingly, given the importance of \emph{medlyn\_slope} on stomatal responses to CO$_{2}$, we did not find a widespread impact of the Medlyn parameter on the fertilization response. Earlier studies using a less constrained range of \emph{medlyn\_slope} values indicated that it can have a potentially large impact on system state, (results not shown) but that the fertilization response is typically limited. The linear response of the stomatal conductance term to $CO_{2}$ mean that the direct impact of \emph{medlyn\_slope} is likely to be restricted to changes in the pre-fertilization state.

The impact of \emph{cn\_flex\_c} was small compared to the more dominant parameters. Higher values of \emph{cn\_flex\_c} spend less carbon on N uptake when tissue N is lower than the target amount, allowing N ratios to vary more. Impacts of this parameter are thus manifested through their impact on $N_{leaf}$ (Figure SI **).  Smaller values of \emph{cn\_flex\_c} corresponded in all cases to smaller variations in $N_{leaf}$ (values closer to 1.0) under both CO$_{2}$ and N fertilization (For N deposition, $N_{leaf}$ always increases and for CO$_{2}$ it always declines.) Lower values of \emph{cn\_flex\_c} also corresponded to larger CO$_{2}$ stimulation of GPP, illustrating the direct impact of this parameter, as leaves maintain higher photosynthetic capacity under more N limiting conditions. The impact on NPP and LAI was less clear, given the corresponding impacts of higher respiration both for maintenance and for N uptake. At CAX, NWR and HVF,  the highest values of \emph{cn\_flex\_c} \emph{also} had a higher response to CO$_{2}$ than the default. The reasons for this discontinuity are unclear, however, the impact of \emph{cn\_flex\_c} in the CLM5 code is subject to at least two 'caps', (whereby plants cannot spend less than 0.5 or more than 1.0 of the carbon that would retain the target LAI), potentially leading to complex feedbacks once these caps are reached.

@@

To interrogate this further, we ran an entire separate set of simulations with higher allocation to leaves (generated from lower stem allocation parameters). These simulations indicated that, with the resulting higher, and more consistent with observed, LAI values, the model was less sensitive to both CO$_{2}$ and N deposition. In the default parameterization CO$_{2}$ and N fertilization increased LAI, amplifying these magnitude of ecosystem responses to these perturbations. For example, at Harvard Forest, the high leaf allocation simulation (LAI of 4.6, vs 2.6 for the default parameters) had a lower CO$_{2}$ response (1.18, 1.17 and 1.22 for GPP, NPP and LAI, respectively, in the high LAI case vs. 1.43, 1.66, 1.47 in the control) and a lower N response as well (1.05, 1.20 and 1.15 for GPP, NPP, and LAI, respectively, in the high LAI case vs. 1.1, 1.26, 1.21 for the control). Similar results were found at all other sites.  Thus, by virtue of their low LAI, the overall fertilization rates at these sites may be substantially larger than those in the model driven with reanalysis meteorology.

\section{Discussion}
The overall responses of the land surface components of Earth system models to CO$_{2}$ fertilization are the subject of intense investigations, and vast quantities of computing time will be expended on the single release version of each model, under the auspices of the various CMIP6 model inter-comparison projects and associated activities (\cite{meehl2014}, \cite{eyring2016}). Throughout the CMIP activities, the carbon-cycle feedbacks reported by each modeling group will be from a single default instance of the parameter space.

At the time of writing, we do not yet know the climate feedback strength of the default version of CLM5/CESM2 compared to its predecessors, but in offline simulations forced with a data atmosphere, CLM5 shows a stronger response to elevated CO$_{2}$ than previous versions of the model (CLM4 and CLM4.5, Wieder et al. 2018). Given the dominant impact of N fixation on the model fertilization responses, it seems likely that the capacity for the model to fix addition N in response to CO$_{2}$ fertilization is at least partly responsible for the change in the global responses, and also that any changes in this parameter would be likely to have large impacts on this result.  Intuitively, it seems likely that an ecosystem that could more rapidly shift a large fraction of resources to N uptake would

@@

  \begin{figure}[h]
     \includegraphics[width=1.2\textwidth]{matlab/figures/MAY19jp_CNtime_defpft.png}
     \caption{Fertilization responses of the default model though time since the (step-change) in CO$_{2}$ (top row) and Nitrogen deposition (bottom row) for CAX (blue), NWR (red), HVF (yellow), MET (purple).}
     \label{timescales}
 \end{figure}
